### 2022-05-14 paper.tex
first commit

\documentclass[review,12pt]{elsarticle}
\usepackage[left=3.2cm, right=3.2cm, top=3.2cm, bottom=3.2cm]{geometry}

\usepackage{lineno, hyperref}
\modulolinenumbers[5]

% \journal{Theoretical Economics}
\usepackage{graphicx}
% \usepackage[abbr]{harvard}
\usepackage{amsmath}
% \usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{braket}
\usepackage{threeparttable}
\usepackage[table]{xcolor}

% Solution to fix compiler warning https://tex.stackexchange.com/a/515199
\usepackage{regexpatch}
\makeatletter
\regexpatchcmd\ps@pprintTitle
  {\cE\}\ \cE\}}{\cE\}\cE\}}
  {}{\FailedToPatch}
\makeatother

% \RequirePackage[colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue,pagebackref]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Uncomment next line to change            %%
%% the type of equation numbering           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\numberwithin{equation}{section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Assumption, Axiom, Claim, Corollary, %%
%% Lemma, Theorem, Proposition, Hypothesis, %%
%% Fact                                     %%
%% use \theoremstyle{plain}                 %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \theoremstyle{plain}
\newtheorem{axiom}{Axiom}
\newtheorem{postulate}{Postulate}
\newtheorem{claim}[axiom]{Claim}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newdefinition{remark}{Remark}
\newdefinition{definition}{Definition}
\newproof{proof}{Proof}
% \newtheorem*{fact}{Fact}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Definition, Example, Remark,         %%
%% Notation, Property                       %%
%% use \theoremstyle{remark}                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \theoremstyle{remark}
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem*{example}{Example}

\newcommand{\proofofref}{}
\newproof{zproofof}{Proof of \proofofref}
\newenvironment{proofof}[1]
 {\renewcommand{\proofofref}{#1}\zproofof}
 {\endzproofof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Local Definitions definitions:           %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\DeclareMathOperator{\Cov}{\mathrm{Cov}}
\DeclareMathOperator{\Tr}{\mathrm{Tr}}
\newcolumntype{L}{>{\centering\arraybackslash}p{0.085\linewidth}}

\begin{document}

\begin{frontmatter}

  \title{Algorithm for Calculating Lightning Node Fees}

  %   %% Group authors per affiliation:
  \author[1]{Cal Abel\texorpdfstring{\footnote{Declarations of interest: none}}}\corref{cor1}
  \address[1]{2600 Century Parkway, Ste. 100, Atlanta, GA, USA}

  %   %% include affiliations in footnotes:
  \author[1]{Skylane Engineering, LLC}
  \ead[url]{https://skylaneengineering.com}

  \cortext[cor1]{Corresponding author}
  \ead{crabel@skylaneengineering.com}

  %   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %   %% Abstract                                 %%
  %   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{abstract}

  \end{abstract}

  %   \begin{keyword}

  %   \end{keyword}

\end{frontmatter}

\linenumbers

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Introduction                                                                                %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
The objective of this paper is to lay out the structure of an algorithmic fee structure for routing channel in the Lightning Network.
With current technology, channels are limited in their ability to dynamically assess fees based on channel balance.
What we propose is a method by which a node will use the fee structure to become self balancing, and charging a higher fee for transactions that are routed that take a channel out of balance.

To do this wee need to first formally define a metric for channel balancing.
This is done by treating the sats locked up in a channel as a microcanonical ensemble.
This will provide us with a combinatorial definition of entropy, which when maximized will define a channel being balanced.

The overall approach of this method is to set up a fee structure using the equivalent of the Helmholtz potential.
We will look at cases where a node operator may choose to entangle some channels based on a desire to ensure certain channels have a higher entropy than others based on transaction traffic.

\section{Defining the Ensemble}
We begin by considering a node in isolation from the rest of the network.
When a node is created traditionally created it is an entirely self funded channel.

\subsection{Ensemble of 1 Satoshi}
Let's look at a single sat channel.
The sat can be in either the inbound state, $\Ket{I}$, or an outbound state, $\Ket{O}$.
Because the sat exists classically a pure state cannot exist as a superposition,
\begin{equation}
  \Ket{\psi} = c_I\Ket{I} + c_O\Ket{O} \; \left\{c_I,c_O\right\} \in \left\{0,1\right\} \textrm{ and } c_I \neq c_O \nonumber
\end{equation}

For this single satoshi we will define its potential relative to the direction that it can be spent.
Thus the sat in $\Ket{I}$ will have a $(+)$ potential and in $\Ket{O}$ a $(-)$ potential.
Because the magnitude value of the satoshi is symmetric we can define the the Hamiltonian as,
\begin{equation}
  \hat{H} =   \left[ {\begin{array}{cc}
          u_0 & 0    \\
          0   & -u_0 \\
        \end{array} } \right].\nonumber
\end{equation}
Thus,
\begin{eqnarray}
  \hat{H}\Ket{I} & = & u_0\Ket{I} \textrm{ and} \nonumber\\
  \hat{H}\Ket{O} & = & -u_0\Ket{O}. \nonumber
\end{eqnarray}

\subsection{Ensemble of Many Satoshi}
We define a channel with $M$ total satoshi and describe the occupancy of the inbound state as $m_I$ and the outbound state as $m_O$.
With the channel so defined we want to to now count the number of different ways that the channel, its multiplicity which is
\begin{equation}
  W = \frac{M!}{m_I!m_O!}. \nonumber
\end{equation}
And define its logarithm as,
\begin{equation}
  S = \log W. \nonumber
\end{equation}
Which for large $M$, and applying the Stirling approximation simplifies to,
\begin{equation}
  S = \sum_j m_j \log\frac{M}{m_j}.\label{eq:1}
\end{equation}
Equation \ref{eq:1} represents the measure of the channels size and, when maximized its balance

\subsection{The Canonical Ensemble}
Because a channel is fixed in its size, the total number of satoshi are invariant with time.
If we bring the channel onto the network where other channels can access it to route payments, we can consider it as an ensemble of a fixed size brought into contact with a reservoir.
This is the canonical ensemble, whose most probable density matrix is,
\begin{equation}
  \hat{\rho} = \frac{1}{Z}e^{-\beta M \left\lvert \hat{H} \right\rvert}. \nonumber
\end{equation}
Where $Z$ is the normalization constant and is also called the partition function.

\subsection{Perturbations form Equilibrium}
The equilibrium state occurs when the channel is balanced thus the probability of finding a satoshi in either the inbound or outbound channels is $p_I=p_O=1/2$.
This isn't very interesting mathematically so we will consider small perturbations around the equilibrium state.
We recall for the canonical ensemble the Helmholtz potential is,
\begin{equation}
  F \equiv - \frac{1}{\beta}\log Z = \Braket{U} - TS. \nonumber
\end{equation}
Where, $T \equiv 1/\beta$.

For a small perturbation,
\begin{equation}
  F + \delta F = \Braket{U} + \left\lvert \delta \Braket{U} \right\rvert - T(S + \delta S).\footnote{
    We take the absolute value of the perturbation because our definition of the sign of the Hamiltonian eigenvectors was entirely arbitrary.
  } \nonumber
\end{equation}
Which reduces to,
\begin{equation}
  \delta F =  \left\lvert \delta \Braket{U} \right\rvert - T \delta S. \nonumber
\end{equation}

Finally if we take the difference between two perturbations,
\begin{equation}
  \Delta F =  \left\lvert \Delta \Braket{U} \right\rvert - T \Delta S. \label{eq:2}
\end{equation}

We take equation \ref{eq:2} as the difference in the potential between the two near equilibrium channel configurations.
Recalling that the Helmholtz potential represents the amount of available work, we will take this as the available potential for extracting a fee.

The $\left\lvert \Delta \Braket{U} \right\rvert$ represents the traditional channel ppm fee.
However there is an additional fee for moving the channel farther from balance and for bringing it closer to balance, $- T \Delta S$.
It is important to note that the derivation did not include a per transaction fee.

Because equation \ref{eq:2} can be negative, we will exclude the case where the channel operator would pay to balance the channel.
The transaction fee is thus,

\boxed{
  \begin{aligned}
     & \textrm{Fee} = \max(0,\Delta F). \nonumber
  \end{aligned}
}

\section{Measuring Temperature}
Recall in defining the ensemble that we did not consider the time dependent, kinetic, portion of the Hamiltonian.
It is not something that is clearly defined for sats on a lightning channel.

Instead we need to look at how the channel is being used in the entire network.
To do this we need to look at the transaction distribution across the channel or for the node.
This distribution should follow some form of the gamma distribution.
If this provides a reasonable estimation, the rate factor, $\beta$, is the inverse temperature \cite[p.~76]{Gibbs:1902}.

The node operator would then need to determine what the fractional charge for temperature would be.
This in effect becomes an average per transaction fee of the change in the entropy.
It's not clear at this time if the measured channel temperature should be used to assess the fee directly, or if it should be multiplied by some constant.

\section{Entanglement}
A node operator may wish to preferentially encourage or penalize a particular node to which they have a channel.
We can do this by looking expanding our single channel ensemble to that of the composite of the node.
This is done through the node's combined Hamiltonian.

If we take small routing node with three channels, there are $2^3$ possible routing combinations.
What the node operator can do is to entangle a pair of routing combinations by adjusting the corresponding off diagonal component of the Hamiltonian.
This will change the overall fee structure for the channel by correspondingly increasing or decreasing the ppm routing fee.

\section{Conclusion}
Hopefully this will all work...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Acknowledgements                                                                            %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Acknowledgements}

\bibliography{paper.bib}

\end{document}

### 2022-05-14 paper.tex
remove entanglement

@@
   \begin{abstract}
-
+    This paper outlines a simple strategy to create channels on a Lightning Network routing node that will tend toward a balanced configuration.
+    The algorithm is derived from the principles of statistical mechanics and uses the Helmholtz potential to dynamically assess either inbound or outbound fees.
   \end{abstract}
@@
 The overall approach of this method is to set up a fee structure using the equivalent of the Helmholtz potential.
-We will look at cases where a node operator may choose to entangle some channels based on a desire to ensure certain channels have a higher entropy than others based on transaction traffic.
 
@@
 
-\section{Entanglement}
-A node operator may wish to preferentially encourage or penalize a particular node to which they have a channel.
-We can do this by looking expanding our single channel ensemble to that of the composite of the node.
-This is done through the node's combined Hamiltonian.
-
-If we take small routing node with three channels, there are $2^3$ possible routing combinations.
-What the node operator can do is to entangle a pair of routing combinations by adjusting the corresponding off diagonal component of the Hamiltonian.
-This will change the overall fee structure for the channel by correspondingly increasing or decreasing the ppm routing fee.
-
 \section{Conclusion}

### 2022-05-15 paper.tex
tie in with paramagent

@@
 \section{Defining the Ensemble}
-We begin by considering a node in isolation from the rest of the network.
-When a node is created traditionally created it is an entirely self funded channel.
+Ensembles are simply collections of a bunch of different things.
+They are extremely useful in physics as their application allows the physicist to be able to describe the stochastic behavior of the ensemble's objects under a different set of constraints.
+These different constraints create a number of different and corresponding physical potentials.
+Three of the more common ensembles are the microcanonical, canonical, and grand canonical.
 
-\subsection{Ensemble of 1 Satoshi}
-Let's look at a single sat channel.
-The sat can be in either the inbound state, $\Ket{I}$, or an outbound state, $\Ket{O}$.
-Because the sat exists classically a pure state cannot exist as a superposition,
+In the case of the microcanonical ensemble we are seeking to describe a system that is in complete isolation from its surroundings, subject to the constraint of its probabilities summing to unity.
+In the canonical distribution, we add the knowledge of the average energy of the system to the constraint of the microcanonical distribution.
+This results in the Helmholtz potential, which expresses the amount of energy available to do work.
+The grand canonical ensemble adds the constraint of the average occupancy of the energy levels to the canonical ensemble.
+This adds the ability for the ensemble to fluctuate in size and energy.
+The grand canonical ensemble results in the Gibbs potential which is a measure of doing chemical work.
+
+While we are not considering physics, we are applying the statistical tools to the problem at hand.
+As a result, it is important to consider how those tools have been interpreted in another context, to get clues on how to interpret them in the current context.
+
+\subsection{Microcanonical Ensemble}
+Let's look at a single sat channel in isolation from its surroundings.
+Borrowing from quantum mechanics, we use the Dirac notation to express our vector space.
+Because our system is so small we can define an orthonormal basis for our satoshi, with two vectors representing whether the satoshi is in the:
+\begin{itemize}
+  \item inbound state $\Ket{I}$ or the
+  \item outbound state $\Ket{O}$.
+\end{itemize}
+
+Because the sat exists classically, a pure state cannot exist as a superposition,
 \begin{equation}
@@
 
-For this single satoshi we will define its potential relative to the direction that it can be spent.
-Thus the sat in $\Ket{I}$ will have a $(+)$ potential and in $\Ket{O}$ a $(-)$ potential.
-Because the magnitude value of the satoshi is symmetric we can define the the Hamiltonian as,
+Now that we have the definition of vector representation of a satoshi, we need to define its potential relative to the direction that it can be "spent".
+For the sat in $\Ket{I}$, because it can move to the outbound state, it has a $(+)$ potential and conversely for $\Ket{O}$ a $(-)$ potential.
+Because the magnitude of the satoshi's value is symmetric, we can define the the Hamiltonian as,
 \begin{equation}
@@
 \end{equation}
-Thus,
+Using the properties of the orthonormal basis vectors, we have,
 \begin{eqnarray}
@@
 \end{eqnarray}
+Because the Hamiltonian in this case is only describing potentials, it can have negative eigenvalues.
 
-\subsection{Ensemble of Many Satoshi}
+Since the system is in isolation from its surroundings each of the possible basis vectors is equally probable.
+This results in the density matrix,
+\begin{equation}
+  \hat{\rho} = \left[ {\begin{array}{cc}
+          \frac{1}{2} & 0           \\
+          0           & \frac{1}{2} \\
+        \end{array} } \right].\nonumber
+\end{equation}
+This is a similar Hamiltonian to a paramagnet.
+Unlike a paramagnet the definition of the sign of the potential in a lightning channel is arbitrary, and therefore exhibits a symmetry.
+
+We note that under these conditions the expectation of the Hamiltonian is,
+\begin{equation}
+  \Braket{\hat{H}} = \Tr \hat{\rho}\hat{H} = 0. \nonumber
+\end{equation}
+
+\subsection{Canonical Ensemble}
 We define a channel with $M$ total satoshi and describe the occupancy of the inbound state as $m_I$ and the outbound state as $m_O$.
-With the channel so defined we want to to now count the number of different ways that the channel, its multiplicity which is
+With the larger channel defined, we need to count the number of different ways that it can be configured, its multiplicity,
 \begin{equation}
@@
 \end{equation}
-And define its logarithm as,
+The multiplicity, $W$, is the number of equally probable configurations for the ensemble.
+We define the channel's entropy as,
 \begin{equation}
-  S = \log W. \nonumber
+  S \equiv \log W. \nonumber
 \end{equation}
-Which for large $M$, and applying the Stirling approximation simplifies to,
+
+For large $M$ we can simplify the entropy by applying the Stirling approximation,
 \begin{equation}
@@
 \end{equation}
-Equation \ref{eq:1} represents the measure of the channels size and, when maximized its balance
+Equation \ref{eq:1} represents the measure of the channels complexity.
+When $S = \max S$, the system is in its most complex configuration.
+If we were to convert $S$ to a base-2 logarithm, it would represent the information entropy, or the number of binary questions needed to be asked to determine the location of each satoshi.
 
-\subsection{The Canonical Ensemble}
 Because a channel is fixed in its size, the total number of satoshi are invariant with time.
-If we bring the channel onto the network where other channels can access it to route payments, we can consider it as an ensemble of a fixed size brought into contact with a reservoir.
+If we bring the channel onto the network where other channels can access it to route payments, we can consider it as an ensemble of a fixed size brought into contact with a thermal reservoir.
 This is the canonical ensemble, whose most probable density matrix is,
@@
 The equilibrium state occurs when the channel is balanced thus the probability of finding a satoshi in either the inbound or outbound channels is $p_I=p_O=1/2$.
-This isn't very interesting mathematically so we will consider small perturbations around the equilibrium state.
+If the channel had eigenvalues of the Hamiltonian differing in absolute magnitude, while keeping differing signs, the chanel would have a different equilibrium distribution of occupancy.
+
+Next we consider small perturbations around the equilibrium state.
 We recall for the canonical ensemble the Helmholtz potential is,
@@
 \end{equation}
-Where, $T \equiv 1/\beta$.
+Where the temperature is defined as $T \equiv 1/\beta$.
 
@@
 \begin{equation}
-  F + \delta F = \Braket{U} + \left\lvert \delta \Braket{U} \right\rvert - T(S + \delta S).\footnote{
+  F_{\max S} + \delta F = \Braket{U}_{\max S} + \left\lvert \delta \Braket{U} \right\rvert - T(S_{\max S} + \delta S).\;\footnote{
     We take the absolute value of the perturbation because our definition of the sign of the Hamiltonian eigenvectors was entirely arbitrary.
@@
 
+\begin{center}
   \boxed{
     \begin{aligned}
-     & \textrm{Fee} = \max(0,\Delta F). \nonumber
+       & \textrm{Fee} = \max(0,\left\lvert \Delta \Braket{U} \right\rvert - T \Delta S). \nonumber
     \end{aligned}
   }
+\end{center}
 
@@
 
-Instead we need to look at how the channel is being used in the entire network.
-To do this we need to look at the transaction distribution across the channel or for the node.
+Instead we need to look at how the channel/node is being used in the entire network.
+To do this we need to look at the transaction distribution across the channel/node.
 This distribution should follow some form of the gamma distribution.
@@
 
-The node operator would then need to determine what the fractional charge for temperature would be.
+The node operator would then need to determine what the fractional charge for temperature would satisfactory.
 This in effect becomes an average per transaction fee of the change in the entropy.
@@
 \section{Conclusion}
-Hopefully this will all work...
+By adopting a routing fee structure outlined in here, a node signals to the network the state of their channel and how the proposed route impacts the channel to the network through its fee.
+By tying the fee to the state of the node and the relation of the transaction to be routed to the fee, we have a clear
 

### 2022-05-15 paper.tex
Clean things up

@@
     The algorithm is derived from the principles of statistical mechanics and uses the Helmholtz potential to dynamically assess either inbound or outbound fees.
+    If this fee strategy is widely adopted, the network will tend toward a state of equilibrium with the lowest overall fee structure.
+    In the derivation of the fee, the central role that a channel's entropy plays in quantify a channel's status.
   \end{abstract}
@@
 For the sat in $\Ket{I}$, because it can move to the outbound state, it has a $(+)$ potential and conversely for $\Ket{O}$ a $(-)$ potential.
-Because the magnitude of the satoshi's value is symmetric, we can define the the Hamiltonian as,
+Considering a channel that is symmetric, we can define the the Hamiltonian as,
 \begin{equation}
@@
 Where $Z$ is the normalization constant and is also called the partition function.
+This density matrix is different what we would normally have because of the absolute values around the Hamiltonian.
+The reason why the absolute value was added is that the sign in the Hamiltonian was arbitrary and that due to this symmetry the probability of occupancy needs to be independent of the sign.
 
@@
 If the channel had eigenvalues of the Hamiltonian differing in absolute magnitude, while keeping differing signs, the chanel would have a different equilibrium distribution of occupancy.
+The effect that by setting different potentials, the channel can be adjusted to have a different equilibrium distribution of satoshi.
 
-Next we consider small perturbations around the equilibrium state.
+Next we consider small perturbations around the equilibrium state to linearize our model.
 We recall for the canonical ensemble the Helmholtz potential is,
@@
 
-\section{Measuring Temperature}
+\section{Estimating Temperature}
 Recall in defining the ensemble that we did not consider the time dependent, kinetic, portion of the Hamiltonian.
@@
 This in effect becomes an average per transaction fee of the change in the entropy.
-It's not clear at this time if the measured channel temperature should be used to assess the fee directly, or if it should be multiplied by some constant.
+It's not clear at this time if the measured channel/node temperature should be used to assess the fee directly, or if it should be multiplied by some constant.
 
@@
 By adopting a routing fee structure outlined in here, a node signals to the network the state of their channel and how the proposed route impacts the channel to the network through its fee.
-By tying the fee to the state of the node and the relation of the transaction to be routed to the fee, we have a clear
+We have a method that creates an incentive to balance the node, because the fee is directly related to the state of the channel and its relationship relative to the transaction being routed.
+If enough nodes adopt such a strategy for charging fees, the network can reach a state of statistical equilibrium, where the overall entropy of the network is maximized.
+
+Furthermore, it is apparent the importance that a node's channel entropy plays in its ability to route transactions.
+The channel's entropy measures both its capacity, and its overall balance.
+Entropy is a critical measure in channel assessment even if the strategy proposed here for calculating routing fees is not adopted.
 

### 2022-05-23 paper.tex
adjust formalism

--- 2022-05-14 paper.tex
@@
     This paper outlines a simple strategy to create channels on a Lightning Network routing node that will tend toward a balanced configuration.
+    In the derivation of the fee, the central role that a channel's entropy plays in quantify a channel's status is readily apparent.
     The algorithm is derived from the principles of statistical mechanics and uses the Helmholtz potential to dynamically assess either inbound or outbound fees.
+    If this fee strategy is widely adopted, the network will tend toward a state of equilibrium with the lowest overall fee structure.
   \end{abstract}
@@
 \section{Introduction}
-The objective of this paper is to lay out the structure of an algorithmic fee structure for routing channel in the Lightning Network.
+The objective of this paper is to lay out the structure of an algorithmic fee structure for a routing channel on the Lightning Network.
 With current technology, channels are limited in their ability to dynamically assess fees based on channel balance.
@@
 
-To do this wee need to first formally define a metric for channel balancing.
+We will need to provide some general concepts and notation to be able to describe a channel mathematically.
+Channels, when abstractly considered can best be thought of as vector spaces.
+Because of this and due to the physical analogies that we use, we will adopt the Dirac notation for describing complex vector spaces.
+
+Using a classical combinatorics approach to define the possible channel configurations, we propose a metric for a channel's balance.
 This is done by treating the sats locked up in a channel as a microcanonical ensemble.
@@
 
-The overall approach of this method is to set up a fee structure using the equivalent of the Helmholtz potential.
+The overall approach of this method is to set up a fee structure that considers three factors, the transaction size, the balance point, and the net transaction current.
+To develop each of these fees, we will use a variational approach to linearize the fee model to derive expressions for the first fee structures.
+The transaction current is an adaptation of a quantum mechanical model for a paramagnet.
+A node operator will then have four fees to set based upon how they want the channel balanced.
 
-\section{Defining the Ensemble}
-We begin by considering a node in isolation from the rest of the network.
-When a node is created traditionally created it is an entirely self funded channel.
+\section{General Channel Description}
+A channel has two orthogonal states in which a satoshi can be: inbound and outbound.
+We represent the inbound state with the ket vector $\Ket{+}$ and the outbound state with $\Ket{-}$.
+These two vectors provide an orthonormal basis of each satoshi in each channel.
 
-\subsection{Ensemble of 1 Satoshi}
-Let's look at a single sat channel.
-The sat can be in either the inbound state, $\Ket{I}$, or an outbound state, $\Ket{O}$.
-Because the sat exists classically a pure state cannot exist as a superposition,
+Let's take a channel that has been unilaterally funded with $M$ total satoshi.
+Each of the funding satoshi will be in the $\Ket{+}$ state.
+When every satoshi is either all in either $\Ket{+}$ or $\Ket{-}$ we call that a pure state.
+Our example channel has no inbound liquidity because there are no vacant inbound states in which to accept satoshi.
+It can, however, ``spend'' satoshi shifting those satoshi from the inbound state to the outbound state.
+We define the $\Ket{+}\rightarrow\Ket{-}$ transition as having a $(+)$ direction.
+Conversely, the state transition, $\Ket{-}\rightarrow\Ket{+}$, has a $(-)$ direction.
+
+The signs associated with what is positive and what is negative are purely arbitrary.
+As we will later see, the arbitrary nature of the direction follows directly from a symmetry within the channel.
+The symmetry results in the conservation of a channels satoshi.
+Once a channel is funded, its total number of satoshi are fixed until liquidation.
+The only thing that changes with a channel's satoshi is its state.
+
+\subsection{Microcanonical Ensemble}
+We now consider a channel in complete isolation from the remainder of the lightning network.
+This channel forms what is known as a statistical ensemble, where each of the satoshi acts as an independent element within the ensemble.
+Because the satoshi exist classically, we can neglect any quantum interferences and assume that they are orthogonal to each other.
+
+A statistical ensemble that is in complete isolation is called a microcanonical ensemble.
+We define the channels multiplicity as,
 \begin{equation}
-  \Ket{\psi} = c_I\Ket{I} + c_O\Ket{O} \; \left\{c_I,c_O\right\} \in \left\{0,1\right\} \textrm{ and } c_I \neq c_O \nonumber
+  W = \frac{M!}{m_+!m_-!}. \nonumber
 \end{equation}
+Where, $m_+$ is the number of satoshi in the inbound state and $m_-$ is the number of satoshi in the outbound state.
 
-For this single satoshi we will define its potential relative to the direction that it can be spent.
-Thus the sat in $\Ket{I}$ will have a $(+)$ potential and in $\Ket{O}$ a $(-)$ potential.
-Because the magnitude value of the satoshi is symmetric we can define the the Hamiltonian as,
-\begin{equation}
-  \hat{H} =   \left[ {\begin{array}{cc}
-          u_0 & 0    \\
-          0   & -u_0 \\
-        \end{array} } \right].\nonumber
-\end{equation}
-Thus,
-\begin{eqnarray}
-  \hat{H}\Ket{I} & = & u_0\Ket{I} \textrm{ and} \nonumber\\
-  \hat{H}\Ket{O} & = & -u_0\Ket{O}. \nonumber
-\end{eqnarray}
-
-\subsection{Ensemble of Many Satoshi}
-We define a channel with $M$ total satoshi and describe the occupancy of the inbound state as $m_I$ and the outbound state as $m_O$.
-With the channel so defined we want to to now count the number of different ways that the channel, its multiplicity which is
-\begin{equation}
-  W = \frac{M!}{m_I!m_O!}. \nonumber
-\end{equation}
-And define its logarithm as,
+\subsection{Entropy}
+We define a channel's entropy as being,
 \begin{equation}
@@
 \end{equation}
-Which for large $M$, and applying the Stirling approximation simplifies to,
+This is also known as its Boltzmann entropy and is maximal when each state is equally probable.
+The maximum entropy of microcanonical ensemble is $\log 2 M$.
+
+Boltzmann entropy reduces the to the Gibbs entropy by using the Sterling approximation,
 \begin{equation}
-  S = \sum_j m_j \log\frac{M}{m_j}.\label{eq:1}
-\end{equation}
-Equation \ref{eq:1} represents the measure of the channels size and, when maximized its balance
-
-\subsection{The Canonical Ensemble}
-Because a channel is fixed in its size, the total number of satoshi are invariant with time.
-If we bring the channel onto the network where other channels can access it to route payments, we can consider it as an ensemble of a fixed size brought into contact with a reservoir.
-This is the canonical ensemble, whose most probable density matrix is,
-\begin{equation}
-  \hat{\rho} = \frac{1}{Z}e^{-\beta M \left\lvert \hat{H} \right\rvert}. \nonumber
-\end{equation}
-Where $Z$ is the normalization constant and is also called the partition function.
-
-\subsection{Perturbations form Equilibrium}
-The equilibrium state occurs when the channel is balanced thus the probability of finding a satoshi in either the inbound or outbound channels is $p_I=p_O=1/2$.
-This isn't very interesting mathematically so we will consider small perturbations around the equilibrium state.
-We recall for the canonical ensemble the Helmholtz potential is,
-\begin{equation}
-  F \equiv - \frac{1}{\beta}\log Z = \Braket{U} - TS. \nonumber
-\end{equation}
-Where, $T \equiv 1/\beta$.
-
-For a small perturbation,
-\begin{equation}
-  F + \delta F = \Braket{U} + \left\lvert \delta \Braket{U} \right\rvert - T(S + \delta S).\footnote{
-    We take the absolute value of the perturbation because our definition of the sign of the Hamiltonian eigenvectors was entirely arbitrary.
-  } \nonumber
-\end{equation}
-Which reduces to,
-\begin{equation}
-  \delta F =  \left\lvert \delta \Braket{U} \right\rvert - T \delta S. \nonumber
+  S = \sum_{i\in\left\{+,-\right\}} m_i \log \frac{M}{m_i}. \label{eq:1}
 \end{equation}
 
-Finally if we take the difference between two perturbations,
+In our example of a unilaterally funded channel, the one in a pure state, it has an entropy, $S = 0$.
+Because our channel exists classically and without quantum interferences, we cannot describe its state as a linear superposition of its basis vectors.
+A satoshi's state vector, $\Ket{\psi}$ is either,
 \begin{equation}
-  \Delta F =  \left\lvert \Delta \Braket{U} \right\rvert - T \Delta S. \label{eq:2}
+  \Ket{\psi_+} = \begin{pmatrix}
+    1 \\
+    0
+  \end{pmatrix} \textrm{ or }
+  \Ket{\psi_-} = \begin{pmatrix}
+    0 \\
+    1
+  \end{pmatrix} \nonumber
+\end{equation}
+for either the inbound or outbound states respectively.
+
+We can describe the mixed state of any linear combination of satoshi within the channel using a density matrix,
+\begin{equation}
+  \hat{\rho} = \sum_{i\in\left\{+,-\right\}} \frac{m_i}{M}\Ket{\psi_i}\Bra{\psi_i}. \nonumber
+\end{equation}
+From this perspective the Gibbs entropy of equation \ref{eq:1} is the von Neumann entropy, where the diagonal elements of the density matrix are its eigenvalues, $p_i$,
+\begin{equation}
+  \hat{\rho} \Ket{i} = p_i \Ket{i}. \nonumber
+\end{equation}
+And,
+\begin{equation}
+  S = M s  = - M \sum_{i\in\left\{+,-\right\}} p_i \log p_i. \nonumber
 \end{equation}
 
-We take equation \ref{eq:2} as the difference in the potential between the two near equilibrium channel configurations.
-Recalling that the Helmholtz potential represents the amount of available work, we will take this as the available potential for extracting a fee.
+\section{Node Formalism}
+We need to establish a number of definitions and concepts that will help with understanding channels in the context of a routing node.
+This formalism is taken directly from nuclear physics in describing how neutrons interact with matter.
+It is adapted from \cite[pp.~3--6]{Bell:1970}.
 
-The $\left\lvert \Delta \Braket{U} \right\rvert$ represents the traditional channel ppm fee.
-However there is an additional fee for moving the channel farther from balance and for bringing it closer to balance, $- T \Delta S$.
-It is important to note that the derivation did not include a per transaction fee.
+While the only part of the formalism used in this paper is the concept of the \emph{satoshi current}, there are many other techniques that can be used in evaluating the lightning network.
+An interested reader may want to consider specifically Chapter 6 with the variational methods and the adjoint transport equation.
+The adjoint equation is used to describe neutron detectors, which is highly analogous to transaction routing.
 
-Because equation \ref{eq:2} can be negative, we will exclude the case where the channel operator would pay to balance the channel.
-The transaction fee is thus,
+% A quadrature interval is defined by,
+% \begin{equation}
+%   \Delta t_{quad} = sum_i \Delta t_i
+% \end{equation}
+% Where $\Delta t_i = t_i - t_{i -1}$ is the time between transactions through the node.
 
-\boxed{
-  \begin{aligned}
-     & \textrm{Fee} = \max(0,\Delta F). \nonumber
-  \end{aligned}
-}
+A valid route through a node consists of an inbound channel, $\Ket{+_j}$ and an outbound channel, $\Ket{-_k}$.
+The two channels in a particular route form a joint system.
+\begin{equation}
+  \Ket{jk} = \Ket{j}\Ket{k} = \Ket{+_j} \otimes \Ket{-_k}. \nonumber
+\end{equation}
+For the route to be valid one channel has to be $\Ket{+}$ and the other in $\Ket{-}$, thus a route is antisymmetric.
+Because the route is antisymmetric, the permutation operator, $\hat{P}_{jk}$, is defined by,
+\begin{equation}
+  \hat{P}_{j,k} \Ket{jk} = - \Ket{kj}. \nonumber
+\end{equation}
+As a result, from the perspective of the $j^\textrm{th}$ node routes of $\Ket{jk}$ are positive and $\Ket{kj}$ are negative.
 
-\section{Measuring Temperature}
-Recall in defining the ensemble that we did not consider the time dependent, kinetic, portion of the Hamiltonian.
-It is not something that is clearly defined for sats on a lightning channel.
+\begin{definition}[Transaction Velocity]
+  The transaction velocity is the period between transactions along the route $\Ket{jk}$.
+  \begin{equation}
+    \mathbf{v}_i \equiv \frac{1}{\Delta t_i}\Ket{jk}. \nonumber
+  \end{equation}
+\end{definition}
+The base unit of the transaction velocity is transactions per second.
 
-Instead we need to look at how the channel is being used in the entire network.
-To do this we need to look at the transaction distribution across the channel or for the node.
-This distribution should follow some form of the gamma distribution.
-If this provides a reasonable estimation, the rate factor, $\beta$, is the inverse temperature \cite[p.~76]{Gibbs:1902}.
+\begin{definition} [Route Density]
+  A route's density, $M_{j,k,i}$, is the total number of satoshi sent in that route's $i^\textrm{th}$ transaction.
+\end{definition}
+The route density has units of satoshi per transaction.
 
-The node operator would then need to determine what the fractional charge for temperature would be.
-This in effect becomes an average per transaction fee of the change in the entropy.
-It's not clear at this time if the measured channel temperature should be used to assess the fee directly, or if it should be multiplied by some constant.
+The sum of all route densities over all possible routes is the \emph{satoshi density},
+\begin{definition}[Satoshi Density]
+  \begin{eqnarray}
+    m_i & \equiv & \Bra{jk} M_{j,k,i} \Ket{jk} \nonumber \\
+    & = & \sum_{j,k}M_{j,k,i} \nonumber
+  \end{eqnarray}
+\end{definition}
+
+We need to consider the \emph{vector flux} across a particular channel, which is the product of the \emph{transaction velocity} and the \emph{route density},
+\begin{definition}[Vector Flux]
+  \begin{equation}
+    \mathbf{\Phi}_{j,k,i} \equiv M_{j,k,i}\mathbf{v}_{j,k,i}. \nonumber
+  \end{equation}
+\end{definition}
+
+The sum of the vector flux over all possible routes for a node is the \emph{total flux} or \emph{scalar flux},
+\begin{definition}[Total Flux]
+  \begin{equation}
+    \phi_i \equiv \Bra{jk}\mathbf{\Phi}_{j,k,i} = \sum_{j,k} \frac{M_{j,k,i}}{\Delta t_i}. \nonumber
+  \end{equation}
+\end{definition}
+
+We want to consider the rate at which satoshi are crossing a channel in either the inbound or outbound directions.
+The rate of satoshi crossing a channel is called the \emph{satoshi current} whose partial currents are,
+\begin{definition}[Partial Current]
+  \begin{eqnarray}
+    \mathbf{J}_{j,i}^+ & = & \Bra{k}\mathbf{\Phi}_{j,k,i} = \sum_k \frac{M_{j,k,i}}{\Delta t_i}\Ket{j} \nonumber \\
+    \mathbf{J}_{j,i}^- & = & \Bra{k}\mathbf{\Phi}_{k,j,i} = -\sum_k \frac{M_{k,j,i}}{\Delta t_i}\Ket{j} \nonumber
+  \end{eqnarray}
+\end{definition}
+And the total \emph{satoshi current} which is the sum of the partial currents,
+\begin{definition}[Satoshi Current]
+  \begin{equation}
+    \mathbf{J}_{j,i} \equiv \mathbf{J}_{j,i}^+ + \mathbf{J}_{j,i}^- \nonumber
+  \end{equation}
+\end{definition}
+
+\section{The Hamiltonian and the Canonical Ensemble}
+We want to think about how the fees effect the state of a channel, because the fees and the channels connections effect how other nodes see this particular node.
+To do this, we define an observable that called the Hamiltonian.
+This observable acts similarly to the Hamiltonian of classical and quantum physics.
+
+The idea is that the fees determine the ``energy'' level (state occupancy) in the context of the transactions that are being routed through that particular channel.
+Lets take a simple case where the inbound fee is $u_+$ and the outbound fee is $u_-$.
+The Hamiltonian is,
+\begin{equation}
+  \hat{H} = \begin{pmatrix}
+    u_+ & 0   \\
+    0   & u_- \\
+  \end{pmatrix}. \nonumber
+\end{equation}
+
+% The expectation of the Hamiltonian is given by,
+% \begin{eqnarray}
+%   \Braket{H_{l}} & = & \Tr\left[\hat{\rho}\hat{H}\right] \nonumber \\
+%   & = & \frac{m_+}{M}u_+ + \frac{m_-}{M}u_- \nonumber
+% \end{eqnarray}
+% If we perturb around the expectation with small fluctuations of $\delta m_+$,
+% \begin{eqnarray}
+%   \Braket{H_{l}}^* & = & \frac{m_+ + \delta m_+}{M}u_+ + \frac{m_- - \delta m_+}{M}u_- \nonumber \\
+%   & = & \Braket{H} + \left(u_+ - u_-\right)\frac{\delta m_+}{M} \nonumber \\
+%   \delta \Braket{H_{l}} & = & \left(u_+ - u_-\right)\frac{\delta m_+}{M}. \label{eq:2}
+% \end{eqnarray}
+% The sign may seem problematic in equation \ref{eq:2}, however, remember we are considering a channel in an outbound transaction having the positive sign and an inbound transaction having a negative sign.
+
+We want to consider the channel being in statistical equilibrium with the network.
+Because the satoshi are fixed on a particular channel, the appropriate ensemble to consider in this case is the canonical ensemble.
+Given our knowledge of the liquidity Hamiltonian, we wish to consider the case where the channel is in its most probable configuration.
+From statistical mechanics we know that for the canonical ensemble this distribution is,
+\begin{equation}
+  \hat{\rho} = \frac{1}{Z}e^{-\beta M \hat{H}}. \nonumber
+\end{equation}
+Where the canonical partition function is,
+\begin{equation}
+  Z[\beta] = \Tr\left[e^{-\beta M \hat{H}}\right]. \nonumber
+\end{equation}
+
+The process of the channel reaching an equilibrium state is the process of a long random walk, which will favor the state with the lower overall potential.
+We will use this property of the canonical ensemble to create better and more adaptive fee structures.
+
+If the channel's liquidity fee is symmetric, each satoshi is equally probable to be in either one state or the other.
+A good analogy to this is the card game War, where each of the players represents either the inbound or the outbound state, and the passing of the pile of cards is a transaction.
+Eventually, one of the states will loose liquidity and have no cards/sats.
+
+Having an asymmetric Hamiltonian allows one of the players to stack the deck slightly in their favor.
+What we will do is to tie the Hamiltonian to the channels state.
+So, as a channel is more out of balance or is being consistently ``pushed'' toward a certain state the fee will increase, to make that particular state statistically less favorable.
+
+We can also use the asymmetric fees to set a desired channel equilibrium state.
+
+\section{Fee Design}
+For the sake of having a standard reference, we will denote an inbound transaction as $\delta m_+$.
+If the channel's inbound state occupancy increases, $\delta m_+ > 0$.
+Similarly, an outbound transaction is $\delta m_-$
+
+\subsection{Liquidity Fee}
+We begin by setting the desired channel equilibrium distribution.
+This is done with the traditionally constant liquidity fee.
+We make one modification to this in that we allow a channel to have a different inbound or outbound fee.
+This will result in the channel being canonically distributed based upon the balancing fee.
+The channel operator would select fixed $u_+$ and $u_-$ such that,
+\begin{equation}
+  \hat{H}_{liq} = \begin{pmatrix}
+    u_+ & 0   \\
+    0   & u_- \\
+  \end{pmatrix}. \nonumber
+\end{equation}
+
+\subsection{Balance Fee}
+If we consider just the entropy of the channel configuration, we are considering only the case of the microcanonical ensemble.
+We need to look at the entropy of the channel's ``energy'', and perturb it with a small transaction.
+\begin{eqnarray}
+  S & = & \frac{2u_+}{u_+ + u_-} m_+ \log \frac{U}{u_+ m_+} + \frac{2u_-}{u_+ + u_-} m_- \log \frac{U}{u_- m_-} \nonumber \\
+  S^* & = & \frac{2u_+}{u_+ + u_-} (m_+ + \delta m_+) \log \frac{U}{u_+ (m_+ + \delta m_+)} \nonumber\\
+  & + & \frac{2u_-}{u_+ + u_-} (m_- - \delta m_+) \log \frac{U}{u_- (m_- - \delta m_+)}  \nonumber \\
+  & \approx & S + \delta m_+ \left( \frac{2u_+}{u_+ + u_-}\log \left(1 + \frac{u_- m_-}{u_+ m_+}\right) - \frac{2u_-}{u_+ + u_-} \log \left(1 + \frac{u_+ m_+}{u_- m_-}\right) \right) \nonumber \\
+  \delta s & = & \frac{2u_+}{u_+ + u_-}\log \left(1 + \frac{u_- m_-}{u_+ m_+}\right) - \frac{2u_-}{u_+ + u_-} \log \left(1 + \frac{u_+ m_+}{u_- m_-}\right) \nonumber
+\end{eqnarray}
+
+We add a balance fee $\tau$ and have the balance portion of the fee Hamiltonian as,
+\begin{equation}
+  \hat{H}_{bal} = \begin{pmatrix}
+    -\tau \delta s & 0              \\
+    0              & -\tau \delta s \\
+  \end{pmatrix}. \nonumber
+\end{equation}
+
+\subsection{Current Fee}
+The final fee that we want to consider is the current fee which uses the satoshi current defined earlier.
+The current across a particular channel acts as a force ``pushing'' the channel's satoshi in a particular direction.
+This is mathematically analogous to what a magnetic field does to electrons in a paramagnet.
+
+We define the current fee as $\mu$, this is equivalent to the magnetic moment of a particle and use the satoshi current to replace the field vector.
+\begin{equation}
+  \hat{H}_{J} = \begin{pmatrix}
+    \mu \,\Bra{j}\mathbf{J}_j & 0                          \\
+    0                         & -\mu \,\Bra{j}\mathbf{J}_j \\
+  \end{pmatrix}. \nonumber
+\end{equation}
+
+Because the satoshi current is described for an instantaneous current of a transaction, the current needs to be aggregated over some quadrature interval, e.g.
+the last 10 transactions for each of the inbound and outbound routes.
+The fee is then applied to the current that is either positive or negative.
+
+\subsection{Combined Fee}
+The combined channel fee is simply the sum of the three sub fees.
+\begin{equation}
+  \hat{F}_j = \begin{pmatrix}
+    u_{j,+} -\tau_j \delta s_j + \mu_j \,\Bra{j}\mathbf{J}_j & 0                                                         \\
+    0                                                        & u_{j,-} -\tau_j \delta s_j -  \mu_j \,\Bra{j}\mathbf{J}_j \\
+  \end{pmatrix}. \label{eq:2}
+\end{equation}
+
+We represent a transaction as,
+\begin{equation}
+  \delta \hat{m} = \delta m \begin{pmatrix}
+    p_+ & 0   \\
+    0   & p_- \\
+  \end{pmatrix}. \nonumber
+\end{equation}
+Because a transaction is either inbound $(+)$ or outbound $(-)$, it is always a pure state, either $\left\{p_+, p_-\right\} = \left\{1, 0\right\}$ or $\left\{p_+, p_-\right\} = \left\{0, 1\right\}$.
+
+Thus the routing fee over route $\Ket{jk}$ is,
+\begin{equation}
+  Fee = \max\left(0,\Tr\left[ \delta \hat{m}_+ \hat{F}_j + \delta \hat{m}_- \hat{F}_k \right]\right). \label{eq:3}
+\end{equation}
+Putting a floor of zero prevents the node operator from paying to route the transaction in the case where the fee potential met the right conditions to be negative.
+It would make sense for a node operator to pay to balance a pair of channels, as they do this during channel rebalancing.
+I just don't know if that is technically possible with the lightning network.
 
 \section{Conclusion}
-Hopefully this will all work...
+By adopting a routing fee structure outlined in here, a node signals to the network the state of their channel and how the proposed route impacts the channel to the network through its fee.
+We have a method that creates an incentive to balance the node, because the fee is directly related to the state of the channel and its relationship relative to the transaction being routed.
+If enough nodes adopt such a strategy for charging fees, the network can reach a state of statistical equilibrium, where the overall entropy of the network is maximized.
+
+Furthermore, it is apparent the importance that a node's channel entropy plays in its ability to route transactions.
+The channel's entropy measures both its capacity, and its overall balance.
+Entropy is a critical measure in channel assessment even if the strategy proposed here for calculating routing fees is not adopted.
 
